Direct Entropy Measurement in a Mesoscopic Quantum System

### 2018-02-02 paper/draft.tex
heaps of edits

% print and edit
% edit figs 3 and 4 text
% edit conclusion

@@

The mesoscopic circuit shown in Fig.~\ref{fig:fig1}a realizes an electron reservoir (central region) in thermal equilibrium with a few-electron quantum dot coupled to its right side, whose occupation is measured using an adjacent charge sensing quantum point contact\cite{Staring2007, Frolov2009, Thierschmann2015}. The number of electrons on the dot can be tuned using the gate voltage, $V_G$, to control the chemical potential.  An electron is added to an $N$ electron dot when the chemical potential to add an electron, $\mu_{N+1}$, drops below the Fermi level of the reservoir, $E_F$.  Figure 1b illustrates a such a transition, showing the expected step in charge sensor conductance as a function of $V_G$ thermally broadened by the reservoir temperature $T_{res}$. The gate voltage corresponding to midpoint of the transition, $V_0$, marks the chemical potential where the probabilities of finding $N$ or $N+1$ electrons on the dot are equal.

Any entropy change from the $N$ to $N+1$ ground states of the dot should shift the midpoint of the transition, $V_0$, by changing the chemical potential on the dot as in Eq.~\ref{eqn:max}.  Detecting the shift in $V_0$ with respect to temperature (${\propto}\frac{\partial \mu}{\partial T}$), amounts to a measurement of the entropy difference $dS$ between the $N$ to $N+1$ ground states ($dN=1$).  In practice, the shift in $V_0$ with temperature is measured by oscillating $T_{res}$ and monitoring any resulting changes in $g_{sens}$ with a lockin amplifier.  If the transition is broadened but not shifted due to heating ($dS=0$), an antisymmetric lineshape is expected from this lockin measurement.  If the transition is not only broadened but also shifted ($dS\neq0$), the change in $g_{sens}$ due to heating will asymmetric, as shown in Fig.~\ref{fig:fig1}c, thus offering a clearly distinguishable signature of entropy for this single charge measurement.

An explanation based on detailed balance is helpful in understanding how the shift in $\mu_{N+1}$ with respect to $E_F$ upon heating relates to entropy.   In general, electrons tunnel back and forth between the dot and reservoir as long as $\mu_{N+1}$ is within thermal broadening of Fermi level, that is, as long as there are available states in both the dot and the reservoir.  The tunnel rates into and out of the dot, $\Gamma_{in}=\Gamma_{N\rightarrow N+1}$ and $\Gamma_{out}=\Gamma_{N+1\rightarrow N}$, depend on degeneracy of the $N$ and $N+1$ levels.  Consider the filling of the first electron state at zero magnetic field, a state with a single, unpaired spin-$\frac{1}{2}$ and therefore a degeneracy of 2. An electron tunneling into this state has both spin-up and -down states available.  Tunneling off, the electron (whether spin up or down) must tunnel into reservoir states with the spin, that is, only half of the reservoir states are available to it. As a result, $\Gamma_{in} = 2\Gamma_{out}$ when $\mu_{N+1}=E_F$. Extending this logic to a general case gives the same results as derived from detailed balance \cite{Gustavsson2009}, 
%
\begin{align}
	\Gamma_{in} &=  g_{N+1} \Gamma f(E_F - \mu_{N+1}) \label{eqn:rates}\\
	\Gamma_{out} &= g_{N} \Gamma [1 - f(E_F - \mu_{N+1})] \nonumber
\end{align}
%
where $\Gamma$ is a coupling constant, $f$ is the Fermi function, and $g_{N/N+1}$ are the degeneracies of the $N$ and $N+1$ electron states.

The midpoint of the transition between the $N$ and $N+1$ electron states, measured by the charge sensor, occurs when the dot is occupied half the time, that is, when $\Gamma_{in} = \Gamma_{out}$. From Eqs.~\ref{eqn:ginout}, $\Gamma_{in} = \Gamma_{out}$ occurs when $\mu_{N+1} = E_F$ as long as $g_{N}=g_{N+1}$, but when $g_{N}<>g_{N+1}$ the midpoint of the transition shifts away from $\mu_{N+1}=E_F$ by an amount that depends on the broadening of the Fermi function and therefore on $T$.   Thus, a measurement of how the midpoint of the transition ($V_0$) shifts with T allows a determination of the ratio $g_{N+1}/g_{N}$, which itself is related to $dS$ through the Boltzmann entropy formula, $S=k_B ln(g)$.  The relationship between degeneracies and tunnel rates was recently explored in a time-resolved measurement \cite{Hofmann2016}.  The thermal technique presented here extends the measurement of entropy (degeneracy) to a wider set of applications, where the tunneling process may not be convenient or even possible to observe directly.

@@

%%% device setup %%%
Using the gates on the right hand side of Fig~\ref{fig:fig1}a, the quantum dot was tuned such that it was tunnel coupled to the reservoir on the source side with the drain side closed. The charge sensing QPC was tuned to a conductance of ${\sim}e^2/h$, where it was found to be most sensitive as a charge detector. A small DC voltage bias was applied across the QPC and the measured current was used to calculate the sensor conductance. The reservoir temperature was controlled either by varying the mixing chamber temperature or by Joule heating through the QPC on the left side of the channel. In the latter case, the QPC was fixed at \SI{20}{\kilo\ohm} and biased with an AC or DC current source. The Joule heating setup gives fast control of the temperature (up to \SI{100}{\hertz}) while keeping the voltage drop across the reservoir negligibly small.

%%% figs 2a and b %%%
Fig.~\ref{fig:fig2}a shows the charge sensor conductance as the first electron is added to the dot by tuning $V_G$, at two different temperatures. The higher temperature transition curve is broadened by the increased temperature and shifted left due to the change in entropy (degeneracy)  between the empty and singly-occupied states. Assuming the source barrier is sufficiently large, $\Gamma_{in/out}$ are broadened only by the temperature of the reservoir and the conductance takes the form,
%
\begin{align}
\label{eqn:g}
        g_{sens}(V,\Theta) &= G_0 \tanh\left(\frac{V-V_0(\Theta)}{2\Theta}\right)  \\
                        &\quad + G_1\left[V-V_0(\Theta)\right] + G_2 \nonumber
\end{align}
%
where $\Theta = \frac{k_B T}{\alpha e}$, $\alpha$ is the lever arm for the quantum dot plunger gate, and $G_1$ is determined by the cross capacitance between the sensor gates and the plunger gate. Fits to this equation are seen in addition to the sensor data in Fig.~\ref{fig:fig2}a. From the fits the transition width in mV, $\Theta$, can be extracted. Fig.~\ref{fig:fig2}b $\Theta$ is plotted as a function of mixing chamber temperature, $T_{MC}$. From this plot it is clear that Eq.~\ref{eqn:g} holds when $T_{MC}\geq$\SI{100}{\milli\kelvin}. For this reason, the mixing chamber temperature remains mixed at \SI{100}{\milli\kelvin} for the remainder of this work. Fitting high temperature data in Fig.~\ref{fig:fig2} to the definition of $\Theta$ (with $T=T_{MC}$) yields $\alpha$, making it possible to convert from $V_G$ to $\mu$ directly.

%%% figs 2c and d%%%
Data in Fig.~\ref{fig:fig2}c, and the corresponding fits, illustrate how entropy was measured at a single charge transition. Fluctuations in $g_{sens}$, $dg_{sens}$, were measured as $T_{electron}$ in the reservoir was varied using Joule heating with an AC current source ($f_{Heat}$\SI{48.7}{\hertz}). The change in $T_{electron}$ of the reservoir is given by $dT \sim I^{AC}_{Heat} \sim \sin^2(2\omega t)$. A calibration for $dT$, using $\alpha$ as measured in Fig.~\ref{fig:fig2}b, can be seen in Fig.~\ref{fig:fig2}c. To avoid thermally accessing excited states of the dot it must be true that $k_B dT \ll \Delta$ where $\Delta$ is the level spacing between the ground and first excited states; which holds for all transitions investigated.

@@

%%% fig 2e %%%
Note that $\delta$ is a fit parameter independent of $dT$ and $\alpha$. Fig.~\ref{fig:fig2}e shows that $dS$, as determined from fits to $dg_{sens}$ data, remains constant over a range of $dT$ ($I^{AC}_{Heat}$). The vertical scale on the right side of Fig. \ref{fig:fig2}e, $\delta$ in units of gate voltage per unit temperature, illustrates the necessity of the lockin measurement; device stability and small signals made it difficult for $delta$ to be determined directly from $g_{sens}$ at different $T$.

@@

Fig. \ref{fig:fig3}c shows $dS$ as a function of parallel field for the 0-1 (and 2-3) electron transitions. The spin-$\frac{1}{2}$ degeneracy of the 1-electron ground state is lifted as the field increases to \SI{\pm5}{\tesla}.  As the Zeeman splitting, measured directly in Fig. \ref{fig:fig2}b, lifts the spin degeneracy, the entropy of the 1-electron state goes to zero. The 2-3 electron transition behaves similarly at low fields.  The 2-electron ground state is the spin singlet, and the 3-electron ground state is same singlet with an additional unpaired electron, with degeneracy of 2, as in the 1-electron case. Below \SI{5}{\tesla} $g\mu_B B$ remains the largest energy scale, and no 3-electron excited states need be considered.

To understand the behavior of $dS$ as a function of magnetic field, Eq. \ref{eqn:S} can be rewritten as,
%
\begin{align}
\label{eqn:Sp}
        S &= k_B \sum_{i=\pm} p_{i}(B, T) \ln{ p_{i}(B,T) }
\end{align}
%
where $p_{i}(B, T)$ is the probabilty of the system being in the i$^{th}$ microstate at a given field and temperature. The 1 and 3-electron ground states have only a spin-$\frac{1}{2}$ degree of freedom at low field and $p_{i}(B, T)$ can be written as a two-level system,
%
\begin{align}
\label{eqn:p}
        p_{\pm}(B, T) &= \frac{1}{1+ e^{\mp \frac{g\mu_B B}{k_B T}}}
\end{align}
%
where $\pm$ represent the spin up/down states of the unpaired electron.

The dashed and solid lines in Fig \ref{fig:fig3}c show fits to $dS$ using Eqs. \ref{eqn:Sp} and \ref{eqn:p} for the 0-1 and 2-3 electron transitions, respectively. These fits allow for peak height, g-factor, and a vertical offset as free parameters; $T$ is fixed at the mixing chamber temperature. From these fits, $dS=0.92 \ln{2}$ and $g=-0.33$. The peak height, corresponding to $dS(B=0)$, deviates from the expected value because it is convoluted with the vertical offset, caused by the cross-capacitance between plunger and QPC gates. The discrepancy between $g$ and the result in Fig. \ref{fig:fig3}b, can be understood from the change in shape of the quantum dot when tuning between the open and closed drain configurations [citation?].

@@

The low field (\SI{<5}{\tesla}) results in Fig. \ref{fig:fig4}c can be understood as simply the inverse of Fig \ref{fig:fig3}c. The entropy of the 2-electron ground state remains $0$, in the spin-singlet state, and the entropy of the 1-electron ground state goes from $k_B\ln{2}$ to $0$ due to the Zeeman splitting seen in Fig. \ref{fig:fig3}b. At high field, the 1-electron ground state remains non-degenerate, while the 2-electron ground state picks up a degeneracy of 2 at \SI{9}{\tesla} as the singlet and triplet ($T_+$) levels cross. This confirms that the $dS$ signal can be related directly to a tunable degeneracy in the system. The same splittings in the quantum dot conductance data can be seen in Fig. \ref{fig:fig4}b.

The entropy signal for the 1-2 transition can be fit using Eq. \ref{eqn:p} as the probabilities for the 1-electron groundstates and the following probabilities for the 2-electron singlet-triplet states,
%
\begin{align}
\label{eqn:pst}
        p_{S/T}(B, T) &= \frac{1}{1+ e^{\mp \frac{g\mu_B B - \Delta_{ST}}{k_B T}}}
\end{align}
%
where $\Delta_{ST}$ is the singlet-triplet splitting. Looking at the width of the peaks in $dS$, it is clear that the g-factor changes, along with the singlet-triplet splitting, as the field is increased. Fitting the curve yields $g(B{=}0T)=-0.46$, $g(B{=}9T)=-0.31$, $\Delta_{ST}(B{=}9T)$ = \SI{240}{\micro\electronvolt}, $dS(B{=}0T) = 1.04\ln{2}$, and $dS(B{=}9T) = 1.01\ln{2}$. Each of these values is consistent with the dot transport data and the fits in Fig. \ref{fig:fig3}c.

The results presented in this work have confirmed the degeneracy of the few-electron ground states states, as calculated using only the spin degree of freedom. Magnetic field dependence of the entropy can be fit using simple spin statistics.  Additionally, a technique has been demonstrated to measure the entropy of few-particle states in mesoscopic devices. Future work may take advantage of this strategy to probe the statistics of non-trival electron and quasiparticle states.

>>>>

@@

The mesoscopic circuit shown in Fig.~\ref{fig:fig1}a realizes an electron reservoir (central region) in thermal equilibrium with a few-electron quantum dot coupled to its right side, whose occupation is measured using an adjacent charge sensing quantum point contact\cite{Staring2007, Frolov2009, Thierschmann2015}. The number of electrons on the dot can be tuned using the gate voltage, $V_G$, to control the chemical potential.  An electron is added to an $N$ electron dot when the chemical potential to add an electron, $\mu_{N+1}$, drops below the Fermi level of the reservoir, $E_F$.  Figure 1b illustrates a such a transition, showing the expected step in charge sensor conductance as a function of $V_G$ and thermal broadening by the reservoir temperature, $T_{res}$. The gate voltage corresponding to midpoint of the transition, $V_0$, marks the chemical potential where the probabilities of finding $N$ or $N+1$ electrons on the dot are equal.

Any entropy change from the $N$ to $N+1$ ground states of the dot should shift the midpoint of the transition, $V_0$, by changing the chemical potential on the dot as in Eq.~\ref{eqn:max}.  Detecting the shift in $V_0$ with respect to temperature (${\propto}\frac{\partial \mu}{\partial T}$), amounts to a measurement of the entropy difference $dS$ between the $N$ to $N+1$ ground states ($dN=1$).  In practice, the shift in $V_0$ with temperature is measured by oscillating $T_{res}$ and monitoring any resulting changes in $g_{sens}$ with a lockin amplifier.  If the transition is broadened but not shifted due to heating ($dS=0$), an antisymmetric lineshape is expected from this lockin measurement.  If the transition is not only broadened but also shifted ($dS,d\mu\neq0$), the change in $g_{sens}$ due to heating will asymmetric, as shown in Fig.~\ref{fig:fig1}c, thus offering a clearly distinguishable signature of entropy for this single charge measurement.

An explanation based on detailed balance is helpful in understanding how the shift in $\mu_{N+1}$ (with respect to $E_F$) upon heating relates to entropy.   In general, electrons tunnel back and forth between the dot and reservoir as long as $\mu_{N+1}$ is within thermal broadening of Fermi level, that is, as long as there are available states in both the dot and the reservoir.  The tunnel rates into and out of the dot, $\Gamma_{in}=\Gamma_{N\rightarrow N+1}$ and $\Gamma_{out}=\Gamma_{N+1\rightarrow N}$, depend on degeneracy of the $N$ and $N+1$ ground states.  Consider the filling of the first electron state at zero magnetic field, a state with a single, unpaired spin-$\frac{1}{2}$ and therefore a degeneracy of 2. An electron tunneling into this state has both spin-up and -down states available.  Tunneling out, the electron (whether spin up or down) must tunnel into reservoir states with the same spin, that is, only half of the reservoir states are available. As a result, $\Gamma_{in} = 2\Gamma_{out}$ when $\mu_{N+1}=E_F$. Extending this logic to a general case gives the same results as derived from detailed balance \cite{Gustavsson2009}, 
%
\begin{align}
	\Gamma_{in} &=  g_{N+1} \Gamma f(E_F - \mu_{N+1}) \label{eqn:rates}\\
	\Gamma_{out} &= g_{N} \Gamma [1 - f(E_F - \mu_{N+1})] \nonumber
\end{align}
%
where $\Gamma$ is a coupling constant, $f$ is the Fermi function, and $g_{N/N+1}$ are the degeneracies of the $N$ and $N+1$ electron states.

The midpoint of the transition between the $N$ and $N+1$ electron states, measured by the charge sensor, occurs when the dot is occupied half the time, that is, when $\Gamma_{in} = \Gamma_{out}$. From Eq.~\ref{eqn:rates}, $\Gamma_{in} = \Gamma_{out}$ occurs when $\mu_{N+1} = E_F$ as long as $g_{N}=g_{N+1}$. When $g_{N}\neq g_{N+1}$, the midpoint of the transition shifts away from $\mu_{N+1}=E_F$ by an amount that depends on the broadening of the Fermi function, and therefore on $T$.   Thus, a measurement of how the midpoint of the transition ($V_0$) shifts with T allows the ratio $g_{N+1}/g_{N}$ to be determined, which itself is related to $dS$ through the Boltzmann entropy formula, $S_{N}=k_{B} \ln{g_N}$.  The relationship between degeneracies and tunnel rates was recently explored in a time-resolved measurement \cite{Hofmann2016}.  The thermal technique presented here extends the measurement of entropy (degeneracy) to a wider set of applications, where the tunneling process may not be convenient or even possible to observe directly.

@@

%%% device setup %%%
Using the gates on the right side of Fig~\ref{fig:fig1}a, the quantum dot was tuned such that the source was tunnel coupled to the reservoir with the drain closed. The charge sensing QPC was tuned to a conductance of ${\sim}e^2/h$, where conductance was most sensitive to charge on the dot. A small DC voltage bias was applied across the QPC and the measured current was used to calculate the sensor conductance. The reservoir temperature was controlled by varying the mixing chamber temperature or Joule heating through the QPC on the left side of the channel. In the latter case, the QPC was fixed at \SI{20}{\kilo\ohm} and biased with an AC or DC current source. The Joule heating setup gives fast control of the temperature (up to \SI{100}{\hertz}) while keeping the voltage drop across the reservoir negligibly small.

%%% figs 2a and b %%%
Fig.~\ref{fig:fig2}a shows the charge sensor conductance as the first electron is added to the dot by tuning $V_G$, at two different temperatures. The higher temperature transition curve is noticeably broadened by the increased temperature. Assuming the source barrier is sufficiently large, $\Gamma_{in/out}$ are broadened only by the temperature of the reservoir and the conductance takes the form,
%
\begin{align}
\label{eqn:g}
        g_{sens}(V,\Theta) &= G_0 \tanh\left(\frac{V-V_0(\Theta)}{2\Theta}\right)  \\
                        &\quad + G_1\left[V-V_0(\Theta)\right] + G_2 \nonumber
\end{align}
%
where $\Theta = \frac{k_B T}{\alpha e}$, $\alpha$ is the lever arm for the quantum dot plunger gate, and $G_1$ is determined by the cross capacitance between the sensor gates and the plunger gate. The dependence of $V_0$ on $\Theta$ captures any shift in $\mu$ due to a change in entropy (degeneracy) across the transition. Fits to Eq.~\ref{eqn:g} are seen, in addition to the sensor data, in Fig.~\ref{fig:fig2}a. From these fits the transition width, $\Theta$, can be extracted. In Fig.~\ref{fig:fig2}b $\Theta$ is plotted as a function of mixing chamber temperature, $T_{MC}$. From this plot, it is clear that Eq.~\ref{eqn:g} holds when $T_{MC}\gtrsim$ \SI{100}{\milli\kelvin}. For this reason, the mixing chamber temperature remains fixed at \SI{100}{\milli\kelvin} for the remainder of this work. Fitting high temperature data in Fig.~\ref{fig:fig2} to the definition of $\Theta$ (with $T=T_{MC}$) yields $\alpha$, making it possible to convert from $V_G$ to $\mu$.

%%% figs 2c and d%%%
Data in Fig.~\ref{fig:fig2}c, and the corresponding fits, show how $dS$ was measured across a single charge transition. Fluctuations in $g_{sens}$, were measured as $T_{res}$ was varied using Joule heating with an AC current source ($f_{Heat} =$ \SI{48.7}{\hertz}). The change in $T_{res}$ is given by $dT \sim I^{AC}_{Heat} \sim \sin^2(2\omega t)$. A calibration for $dT$, using $\alpha$ as measured in Fig.~\ref{fig:fig2}b, can be seen in Fig.~\ref{fig:fig2}c. To avoid thermally accessing excited states of the dot it must be true that $k_B dT \ll \Delta$ where $\Delta$ is the level spacing between the ground and first excited states; which holds for all transitions investigated.

@@

%%% fig 2e %%%
Note that $\delta$ is a fit parameter independent of $dT$ and $\alpha$. Fig.~\ref{fig:fig2}e shows that $dS$, as determined from fits to $dg_{sens}$ data, remains constant over a range of $dT$ ($I^{AC}_{Heat}$). The vertical scale on the right side of Fig. \ref{fig:fig2}e, $\delta$ in units of gate voltage per unit temperature, illustrates the necessity of the lockin measurement; device stability and small signals made it difficult for $delta$ to be determined directly from $g_{sens}$ at different $T$. Without any additional calibration, it is possible to measure $dS$ directly from the lockin amplifier signal, $dg_{sens}$.

@@

To confirm that the measured signal derives from the electron spin, the behavior of $dS$ is investigated as a function of parallel magnetic field. Fig. \ref{fig:fig3}c shows $dS$ as a function of applied field for the 0-1 and 2-3 electron transitions. The 2-electron ground state is a spin singlet, while the 3-electron ground state is same singlet with an additional unpaired electron, making the field-dependent behavior of these transitions identical where $g \mu_{B} B$ is less than the excited state energies. Zeeman splitting, measured directly in Fig. \ref{fig:fig2}b, lifts the spin degeneracy of the unpaired electrons and the entropy of the 1- and 3- electron ground states goes to zero.

The behavior of $dS$ as a function of magnetic field can be understood using the Gibbs entropy,
%
\begin{align}
\label{eqn:Sp}
        S &= k_B \sum_{i=\pm} p_{i}(B, T) \ln{ p_{i}(B,T) }
\end{align}
%
where $p_{i}(B, T)$ is the probability of the system being in the i$^{th}$ spin state at a given field and temperature.  For states with a single unpaired electron, $p_{i}(B, T)$ can be written as a two-level system,
%
\begin{align}
\label{eqn:p}
        p_{\pm}(B, T) &= \frac{1}{1+ e^{\mp \frac{g\mu_B B}{k_B T}}}
\end{align}
%
where $\pm$ represent the spin up/down states of the unpaired electron. Fits to Eqs. \ref{eqn:Sp} and \ref{eqn:p} are shown in Fig \ref{fig:fig3}c for the 0-1 and 2-3 electron transitions. The fits allow for peak height, g-factor, and a vertical offset as free parameters with $T$ fixed at $T_{MC}$. From these fits, $dS=0.92 \ln{2}$ and $g=-0.33$ for the 0-1 transition.  Discrepancy between $g$ and the result shown in Fig. \ref{fig:fig3}b, is understood from the change in shape of the quantum dot when tuning between the $g_{dot}$ and $g_{sens}$ measurements.

@@

The low field (\SI{<5}{\tesla}) data in Fig. \ref{fig:fig4}c for the 1-2 electron transition is understood as the inverse of Fig \ref{fig:fig3}c. The 2-electron ground state remains in the spin singlet state with zero entropy while 1-electron ground state goes from $k_B\ln{2}$ to $0$ due to Zeeman splitting. At high fields, the 1-electron ground state remains non-degenerate, while the 2-electron ground state becomes two-fold degenerate at \SI{9}{\tesla} where the singlet and triplet ($T_+$) states cross. The singlet-triplet crossing is seen clearly in Fig.~\ref{fig:fig4}b.

The entropy for the 1-2 transition can be fit using Eq. \ref{eqn:p} as the probabilities for the 1-electron states and the following probabilities for the 2-electron singlet-triplet states,
%
\begin{align}
\label{eqn:pst}
        p_{S/T}(B, T) &= \frac{1}{1+ e^{\mp \frac{g\mu_B B - \Delta_{ST}}{k_B T}}}
\end{align}
%
where $\Delta_{ST}$ is the singlet-triplet splitting at zero field. Fitting the curve yields $g(B{=}0T)=-0.46$, $g(B{=}9T)=-0.31$, $\Delta_{ST}(B{=}9T)$ = \SI{240}{\micro\electronvolt}, $dS(B{=}0T) = 1.04\ln{2}$, and $dS(B{=}9T) = 1.01\ln{2}$. Each of these values is consistent with data in Figs.~\ref{fig:fig4}a and b as well as previous GaAs dot measurements [cite why g should change in high field]. This fitting confirms that the $dS$ signal can be related directly to a tunable degeneracy in the system.

%%% conclusion %%%
The results presented demonstrate a novel technique to measure the entropy of few-particle states in mesoscopic devices. The measurement was used to investigate the well-understood spin degeneracy of the few-electron ground states states in a GaAs quantum dot. The entropy signal is confirmed by tuning the entropy as a function of applied magnetic field. Future work may take advantage of this strategy to look for thermodyanmic signatures of non-trival electron and quasiparticle states, eliminating the uncertainty found in typical transport data.
